What We Eval in the Shadows

### 2021-04-14 paper/main.tex
Example of variable lookups

\mypara{Domain Specific Language.} R expressions are built and transformed symbolically and then evalled to a concrete value. For instance,   base R packages \emph{stats} provides two operators for \emph{symbolic differentiation}, \c{D} and \c{deriv}. They support arithmetic operations and functions on real numbers such as \c{sin}.

In \emph{MCMCglmm} (Monte Carl Markov Chain Generalised Linear Mixed Models), a \c{Dtensor} object is created, which contains expressions derived twice (with \c{DD}) and then typically in a list or a data frame.
\begin{lstlisting}
 expr<-expression(beta_1 + time*beta_2+u)
 ...
 mu = data.frame(beta_1=0.5, beta_2=1, time=3, u=2.3)
 ...
 D[i]<-eval(DD(expr, name[unlist(comb.pos[i,])]), mu)
\end{lstlisting}
% See https://github.com/cran/MCMCglmm/blob/cfac9e67ec73a3db2142826faae35e5d2318da31/man/evalDtensor.Rd

>>>>

The dominant site in numbers of calls in CRAN, in the custom object system of \emph{ggplot2}, \c{ggproto}, also evaluates a variable \c{`\_inherit`} in the definition environment of the class. It leads to a large diversity of environments, many being package environment or package environments wrapped in \c{new}.

Package \emph{statnet.common} handles formulas, R object that associates an expression of the form \c{x ~ expr} and a carry-on environment where to evaluate the formula. It evaluates the left-hand side of the formula, which is always a variable, in \c{eval\_lhs.formula}, as \c{eval(object[[2L]],envir=environment(object))} where \c{object} is the formula and \c{environment} returns the formula associated to the object.

%TODO Add a simple example of a slot lookup (very common with match.call)

\mypara{Domain Specific Language.} R expressions are built and transformed symbolically and then evalled to a concrete value. For instance,   base R packages \emph{stats} provides two operators for \emph{symbolic differentiation}, \c{D} and \c{deriv}. They support arithmetic operations and functions on real numbers such as \c{sin}.
In \emph{MCMCglmm} (Monte Carl Markov Chain Generalised Linear Mixed Models), a \c{Dtensor} object is created, which contains expressions derived twice (with \c{DD}) and then typically in a list or a data frame.
\begin{lstlisting}
 expr<-expression(beta_1 + time*beta_2+u)
 ...
 mu = data.frame(beta_1=0.5, beta_2=1, time=3, u=2.3)
 ...
 D[i]<-eval(DD(expr, name[unlist(comb.pos[i,])]), mu)
\end{lstlisting}
% See https://github.com/cran/MCMCglmm/blob/cfac9e67ec73a3db2142826faae35e5d2318da31/man/evalDtensor.Rd
